---
title: "SQL and R"
author: "Nataliya Ferdinand"
date: "2024-09-08"
output: html_document
---

# Introduction

For this project, I selected six popular movies and asked five individuals from my social circle, including friends and family, to rate each movie they had seen on a scale from 1 to 5. I gathered these ratings and observations and stored them in a structured SQL database. To facilitate analysis and further exploration, I then connected to the SQL database and loaded the movie rating data into an R data frame for processing. This workflow highlights my ability to work with relational databases, SQL queries, and data manipulation in R. 

#### **Setting a CRAN mirror before installing the package "DBI"**

When I am trying to install the DBI package in my RMarkdown file, R is attempting to connect to CRAN without specifying which mirror (server) to use. Thus, I need to set a CRAN mirror before installing the package "DBI"

```{r}
options(repos = c(CRAN = "https://cloud.r-project.org"))
```
##### **Installing packages **

I installed "DBI" package that provides an easy way to connect with SQL databases, allowing to obtain, modify, and analyze data in real-time.
Also, I installed "RMySQL" package since I am using MySQL. 

```{r}
install.packages("DBI")
install.packages("RMySQL")
```

#### **Establishing Database Connection**

I need to establish a connection to my SQL database (MySQL) before running queries. 

```{r}
library(DBI)
movie_database<-dbConnect(RMySQL::MySQL(),
               dbname="movie",
               host ="localhost", #or my SQL server address
               port = 3306,
               user = "root",
               password="Matrix%607")

#check if connection is successful
dbListTables(movie_database)
```
```{r}
movies_table<-dbGetQuery(movie_database, "select * from movies")
movies_table
```
#### **Checking all columns of all three tables: users, movies, and users_movies **

I created three tables and I want to see all the columns and the data in each column. 

```{r}
users_table<-dbGetQuery(movie_database, "select * from users")
users_table
```

```{r}
users_movies_table<-dbGetQuery(movie_database, "select * from users_movies")
users_movies_table
```

#### **The SQL query to join the tables**

Since the columns of interest locates in different columns, I need to join them. The columns which I want to get are userid  and username from "Users" table; movieid and title are from "movies" table; and rating column is from "users_movies" table. 

```{r}
query1<-"select u.userid, u.username, t.movieid, t.title, r.rating 
       from movies t JOIN users_movies r 
       ON t.movieid=r.movieid
       JOIN users u ON u.userid=r.userid"
result1<-dbGetQuery(movie_database, query1)
result1
```

#### **Installing "dplyr" package**

I need "dplyr" package for making data manipulation easier. For example, I filter only those observations (rows) that do not have missing data (NA).

```{r}
library(dplyr)
result_clean<-result1%>%filter(!is.na(rating))
result_clean
```

#### **Calculation of average rating per movie**

In order to create a graph that will show an average rating for each movie, I used summarize(), mean() group_by() functions. 

```{r}
# Calculate average rating per movie
avg_rating_per_movie <- result_clean %>%
  group_by(title) %>%
  summarise(avg_rating = mean(rating))
avg_rating_per_movie
```

#### **Creation of a bar plot of average ratings** 

```{r}
# Create a bar plot of average ratings
library(ggplot2)
ggplot(avg_rating_per_movie, aes(x = title, y = avg_rating)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Average Rating per Movie", x = "Movie Title", y = "Average Rating")
```

```{r}
dbDisconnect(movie_database)
```



# Conclusion: 

After analyzing the ratings from all five participants, the movie *Game of Thrones* emerged as the highest-rated, consistently receiving strong feedback. Based on the ratings provided, it secured the top spot with the highest average score, demonstrating its popularity and appeal among the group of respondents. 
